Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-12-21 demystify.tex
updates

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{soul}
\usepackage[most]{tcolorbox}
% \setsecnumdepth{subsection}
% \setcounter{tocdepth}{3}
\usepackage{enumitem}
\usepackage{epigraph}
\usepackage{cite}
\usepackage{caption}
\captionsetup{font=small}
\usepackage{graphicx}
\usepackage{pdfpages}
\usepackage{hyperref}
\usepackage{wrapfig}
\setlength\intextsep{0pt} % remove extra space above and below in-line float

@@

\epigraph{\vspace{-0.2in} Don't make fun of graduate students. They just made a terrible life choice.}{\textsc{The Simpsons}}

@@

\epigraph{How is education supposed to make me feel smarter? Besides, every time I learn something new, it pushes some of the old stuff out of my brain. Remember when I took that home wine making course, and I forgot how to drive?}{\textsc{The Simpsons}}

@@

\epigraph{Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.}{\textsc{The Simpsons}}

@@

@@

\epigraph{Bart: ``Are we there yet? Are we there yet? Are we there yet?''\\
Homer: ``No!''\\
Bart: ``Okay, how about now?''}{\textsc{The Simpsons}}

@@

\epigraph{``Oh... and how is education supposed to make me feel smarter? Besides, every time I learn something new, it pushes some old stuff out of my brain. Remember when I took that home wine-making course and I forgot how to drive?''}{\textsc{The Simpsons}}

@@

\epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{\textsc{The Simpsons}}

@@

\epigraph{\vspace{-0.2in} The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!}{\textsc{The Simpsons}}

@@

\subsection{Schools that offer PhD in CS}

@@

\epigraph{``I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here.''}{\textsc{The Simpsons}}

@@

In most schools I've been at, the application deadline is in December, and adcom meets begins when school starts in mid or end of January. Adcom chair sends out review assignments to adcom members, about 30ish per faculty (\autoref{sec:adcom}). We usually have about 2 weeks to review all applications. As mentioned in~\autoref{sec:adcom-discuss}, adcom members review applications individually and independently. We only discuss when there are disagreements (adcom chair will determine which applications to discuss).

I typically reserve a whole day to review all applications (this is to be fair to all applicants).
On average, I spend about 10--15 minutes reviewing each application (less for clear rejections and more for potential acceptances). While this seems short, it is not that difficult to tell if an application is good or bad.  In fact, this
is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.

@@

I read carefully \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers or generic content. I skim over \emph{CV} and look for publications, research experiences, and notable achievements. I take notes of things like ``published papers in top venues'' or ``gold medal in an international competition''.

I skim weak \emph{SOPs} but read strong ones thoroughly. I note whether the SOP is exciting, research-oriented, standing out, e.g., applicant is minority (\autoref{sec:urm}), and tailored to our program (e.g., if they are familiar with the work of some faculty or have talked to them). These notes are entered into the evaluation system.

Finally, I enter my decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA). I also recommend very strong candidates for the University Presidential fellowship. Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I rarely use them, simply because we want to fully fund our PhD students.

@@

\epigraph{\vspace{-0.2in} I didn’t do it. Nobody saw me do it. You can’t prove anything!}{\textsc{The Simpsons}}

@@

\epigraph{\vspace{-0.2in} I’m not a bad guy! I work hard, and I love my kids. So why should I spend half my Sunday hearing about how I’m going to Hell?}{\textsc{The Simpsons}}

@@

\epigraph{\vspace{-0.2in} Kids, you tried your best and you failed miserably. The lesson is: never try.}{\textsc{The Simpsons}}

@@

\epigraph{\vspace{-0.2in} Ay caramba!}{\textsc{The Simpsons}}

@@

\epigraph{\vspace{-0.2in} The whole damn system is wrong!}{\textsc{The Simpsons}}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{soul}
\usepackage[most]{tcolorbox}
% \setsecnumdepth{subsection}
% \setcounter{tocdepth}{3}
\usepackage{enumitem}
\usepackage{epigraph}
\newcommand{\myepigraphsimpsons}[1]{
\epigraph{\vspace{-0.2in} \emph{#1}}{\textsc{The Simpsons}}
}
\usepackage{cite}
\usepackage{caption}
\captionsetup{font=small}
\usepackage{graphicx}
\usepackage{pdfpages}
\usepackage{hyperref}
\usepackage{wrapfig}
\setlength\intextsep{0pt} % remove extra space above and below in-line float

@@

\myepigraphsimpsons{``Don't make fun of graduate students. They just made a terrible life choice.''}

@@

@@

\myepigraphsimpsons{``Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.''}

@@

\myepigraphsimpsons{``To whom it may concern… D’oh!''}

@@

\myepigraphsimpsons{Bart: ``Are we there yet? Are we there yet? Are we there yet?''\\
Homer: ``No!''\\
Bart: ``Okay, how about now?''}

@@

\myepigraphsimpsons{``Oh... and how is education supposed to make me feel smarter? Besides, every time I learn something new, it pushes some old stuff out of my brain. Remember when I took that home wine-making course and I forgot how to drive?''}

@@

\myepigraphsimpsons{``Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like ... love!''}

@@

\myepigraphsimpsons{``The answer to life's problems aren't at the bottom of a bottle. Heh heh! They're on TV!''}

@@

\subsection{Schools offering PhD in CS}\label{sec:schools-offering-phd}

@@

\myepigraphsimpsons{``I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here.''}

@@

In the schools I've been at, the application deadline is in December, and adcom meets when school starts in mid or end of January. Adcom chair sends out review assignments to adcom members, about 30ish per faculty (\autoref{sec:adcom}). We usually have about 2 weeks to review all applications. As mentioned in~\autoref{sec:adcom-discuss}, adcom members review applications individually and independently. We only discuss when there are disagreements (adcom chair will determine which applications to discuss).

I typically reserve a whole day to review all applications.
On average, I spend about 10--15 minutes reviewing each application (less for clear rejections and more for potential acceptances). While this seems short, it is not that difficult to tell if an application is good or bad.  In fact, this
is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.

@@

I read carefully \emph{strong LoRs} and skim weaker ones, noting either strong letters from well-known professors/researchers and talk about research experiences or weak letters with generic content (e.g., ``student was in my class''). I skim over \emph{CV} and look for publications, research experiences, and notable achievements. I take notes of things like ``published papers in top venues'' or ``gold medal in an international competition''.

I skim weak \emph{SOPs} but read strong ones carefully. I note whether the SOP is exciting, research-oriented, standing out, e.g., applicant is minority (\autoref{sec:urm}), and tailored to our program (e.g., if they are familiar with the work of some faculty or have talked to them). These notes are entered into the evaluation system.

Finally, I enter my decision, which is usually either a rejection or an offer of admission with full funding (e.g., from TA). I also recommend very strong candidates for the University Presidential fellowship., which is a fellowship from GMU that provides funding similar to an RA (\autoref{sec:funding}).
Note that while the system has other decision options, eg., admit without funding, provisional admission (e.g., if they need to take some courses), I rarely use them, simply because we want to fully fund our PhD students.

@@

\myepigraphsimpsons{``I didn’t do it. Nobody saw me do it. You can’t prove anything!''}

@@

\myepigraphsimpsons{``I’m not a bad guy! I work hard, and I love my kids. So why should I spend half my Sunday hearing about how I’m going to Hell?''}

@@

\myepigraphsimpsons{``Kids, you tried your best and you failed miserably. The lesson is: never try.''}

@@

\myepigraphsimpsons{``Ay caramba!''}

@@

\myepigraphsimpsons{``The whole damn system is wrong!''}
